$Z_3$ topological order in the quantum dimer-pentamer model

### 2016-06-04 qdpm_paper.tex
fis

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t!]
    \centering
    \includegraphics[width=0.5\columnwidth]{QDPM_ex_config.pdf}
    \caption{Example of an allowed configuration in the quantum dimer-pentamer model.}
    \label{fig:QDPMex}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tpb]
    \centering
    \includegraphics[width=1.0\columnwidth]{mag_loops_on_orntd_lat.pdf}
    \caption{Three oriented magnetic strings on the square lattice. From left to right we show: i) a
    nontrivial string winding around the system in one of the two possible directions, ii) an open
    magnetic string with two different charged (colored) visons at the end points, iii) a trivial closed
    string. The colors red and blue distinguish the two different sublattices that orient the links.}
    \label{fig:orientedSL}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

 %%%%%%%%%%%%%%%%%
%\begin{figure}[tpb]
%\centering
%\includegraphics[width=0.8\linewidth]{pentamer_moves.pdf}
%\caption{a) Horizontal creation an annihilation, b) horizontal move \textcolor{red}{Owen:fix this move}, c) diagonal moves \textcolor{red}{Owen: Remove (d) as it is related to (c) by rotational symmetry}}
%\label{fig:pentamer_moves}
%\end{figure}
%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
    \begin{tabular}{| c | c | c | c | c | c |}
  	\hline
    $c_1$& $\bar{c}_1$ & $c_2$& $\bar{c}_2$ & $c_2$& $\bar{c}_2$ \\
	\hline
    $\pentmvcreate$& $\pentmvanal$ & $\penmvhrzleft$& $\penmvhrzright$ & $\penmvdiagup$& $\penmvdiagdown$ \\
	\hline
  \end{tabular}
  \caption{Defects where $n_v$ dimers touch a vertex and their associated $U(1)$ charges on the A sub-lattice in both the QDM and QFPLM. Defects on the B sub-lattice have opposite charge.}
  \label{tab:defects}
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

We explicitly demonstrate how the extensive topological degeneracy of the QDM is reduced to a finite degeneracy in the QDPM by computing the $U(1)$ topological winding number $\Phi_X$ histogram. This is shown in \figref{fig:u1_wind_qdpm}, where the different colors represent calculations that were initialized to different winding sectors. We confirm that $\Phi_{X,Y}$ are conserved modulo $3$, leading to $9$-fold topological degeneracy. Notice that $\Phi_X = \pm L/2$ correspond to ordered configurations, and therefore the probability of $\Phi_{XY}$ reaching these extrema vanishes in the RK state.
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[]
    \centering
    \includegraphics[width=1.0\linewidth]{u1_wind_qdpm.pdf}
    \caption{The histogram of the $U(1)$ winding number $\Phi_X$ as measured in the QDPM on a $16\times16$ lattice. Different colors correspond to calculations initialized to different winding sectors. Note that $\Phi_X$ is conserved modulo $3$, resulting in a $9$-fold topological degeneracy.}
    \label{fig:u1_wind_qdpm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}
        \centering
        \includegraphics[width=1.0\columnwidth]{dimer_gry_scale_qdpm.pdf}
        \caption{Gray scale of the dimer correlation function between the dimer in the upper
        left corner (colored red) and all other dimers on a $16\times16$ lattice. The 
         red dimers indicate correlations that are off the scale shown at the
        bottom of the figure. \textcolor{red}{Owen: make the origin dimer a different color that the off-scale dimers}.}
        \label{fig:dimer_gry_scale_qdpm}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \includegraphics[width=1.0\columnwidth]{spacial_cors.pdf}
    \caption{Correlation function between parallel dimers one the same (blue) and different (red) 
    sub-lattice on an $L=32$ lattice. The lines represent least-squares fits to exponential decay. }
    \label{fig:spatial_dmr_cor}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
   \begin{figure}[]
        \centering
        \includegraphics[width=0.75\columnwidth]{qdpm_dmr_vis_HM_16x16.pdf}
        \caption{QDPM dimer structure factor $\gamma=0.5$.}
        \label{fig:spacial_dmr_z3vis_cor}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}[]
        \centering
        \includegraphics[width=1.0\columnwidth]{vis_dmr_qdpm_struc_fac_16x16.pdf}
        \caption{The QDPM dimer structure factor $S_d(k)$ and $Z_3$ vison structure factor $S_v(k)$
        shown along the path in $k$ space connecting the $(0,0)$, to $(\pi,0)$, $(\pi,\pi)$, $(0,0)$
        points for $\gamma=0.5$.}
        \label{fig:vis_dmr_cut}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}[]
        \centering
        \includegraphics[width=1.0\columnwidth]{monomer_cor_log.pdf}
        \caption{Monomer correlations in the QDPM.}
        \label{fig:vis_dmr_cut}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t!]
    \centering
    \includegraphics[width=1.0\columnwidth]{QDPM_ex_config.pdf}
    \caption{Example of an allowed configuration in the quantum dimer-pentamer model.}
    \label{fig:QDPMex}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[tpb]
    \centering
    \includegraphics[width=1.0\columnwidth]{mag_loops_on_orntd_lat.pdf}
    \caption{$\Phi_i$: a nontrivial string winding around the system in one of the two possible
        directions. $\Phi_2$: an open
    magnetic string with two different charged (colored) visons at the end points. $\Phi_3$ a trivial closed
    string. The colors red and blue distinguish the two different sublattices that orient the links.}
    \label{fig:orientedSL}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

 %%%%%%%%%%%%%%%%%
%\begin{figure}[tpb]
%\centering
%\includegraphics[width=0.8\linewidth]{pentamer_moves.pdf}
%\caption{a) Horizontal creation an annihilation, b) horizontal move \textcolor{red}{Owen:fix this move}, c) diagonal moves \textcolor{red}{Owen: Remove (d) as it is related to (c) by rotational symmetry}}
%\label{fig:pentamer_moves}
%\end{figure}
%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\begin{table}
%    \begin{tabular}{| c | c | c | c | c | c |}
%  	\hline
%    $c_1$& $\bar{c}_1$ & $c_2$& $\bar{c}_2$ & $c_2$& $\bar{c}_2$ \\
%	\hline
%    $\pentmvcreate$& $\pentmvanal$ & $\penmvhrzleft$& $\penmvhrzright$ & $\penmvdiagup$& $\penmvdiagdown$ \\
%	\hline
%  \end{tabular}
%  \caption{Defects where $n_v$ dimers touch a vertex and their associated $U(1)$ charges on the A sub-lattice in both the QDM and QFPLM. Defects on the B sub-lattice have opposite charge.}
%  \label{tab:defects}
%\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}
    \begin{tabular}{| l | c | c |}
  	\hline
     & $c$ & $\bar{c}$ \\
	\hline
    1 & $\pentmvcreate$ &  $\pentmvanal$  \\
	\hline
    2 &$\penmvhrzleft$&$\penmvhrzright$ \\
	\hline
    3 &$\penmvdiagup$& $\penmvdiagdown$\\
	\hline
  \end{tabular}
  \caption{Defects where $n_v$ dimers touch a vertex and their associated $U(1)$ charges on the A sub-lattice in both the QDM and QFPLM. Defects on the B sub-lattice have opposite charge.}
  \label{tab:defects}
\end{table}

@@

We explicitly demonstrate how the extensive topological degeneracy of the QDM is reduced to a finite degeneracy in the QDPM by computing the $U(1)$ topological winding number $\Phi_X$ histogram. This is shown in \figref{fig:u1_wind_qdpm}, where the different colors represent calculations that were initialized to different winding sectors. We confirm that $\Phi_{X,Y}$ are conserved modulo $3$, leading to $9$-fold topological degeneracy. Notice that $\Phi_X = \pm L/2$ correspond to ordered configurations, and therefore the probability of $\Phi_{XY}$ reaching these extrema vanishes in the RK state.
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[]
    \centering
    \includegraphics[width=1.0\linewidth]{u1_wind_qdpm.pdf}
    \caption{The histogram of the $U(1)$ winding number $\Phi_X$ as measured in the QDPM on a
    $32\times 32$ lattice. Different colors correspond to calculations initialized to different winding sectors. Note that $\Phi_X$ is conserved modulo $3$, resulting in a $9$-fold topological degeneracy.}
    \label{fig:u1_wind_qdpm}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}
        \centering
        \includegraphics[width=1.0\columnwidth]{dimer_gry_scale_qdpm.pdf}
        \caption{Gray scale of the dimer (left) and $Z_3$ vison (right) correlation functions
            The color blue indicates the dimer (plaquete) from which the correlation is measured. The 
            color red indicate correlations that are off the gray scale used here .}
        \label{fig:dimer_gry_scale_qdpm}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}
    \centering
    \includegraphics[width=1.0\columnwidth]{spacial_cors.pdf}
    \caption{Correlation function between parallel dimers one the same (blue) and different (red) 
    sub-lattice on an $L=32$ lattice. The lines represent least-squares fits to exponential decay. }
    \label{fig:spatial_dmr_cor}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
   \begin{figure}[]
        \centering
        \includegraphics[width=1.0\columnwidth]{qdpm_dmr_vis_HM_16x16.pdf}
        \caption{QDPM dimer structure factor $\gamma=0.5$.}
        \label{fig:spacial_dmr_z3vis_cor}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}[]
        \centering
        \includegraphics[width=1.0\columnwidth]{vis_dmr_qdpm_struc_fac_16x16.pdf}
        \caption{The QDPM dimer structure factor $S_d(k)$ and $Z_3$ vison structure factor $S_v(k)$
        shown along the path in $k$ space connecting the $(0,0)$, to $(\pi,0)$, $(\pi,\pi)$, $(0,0)$
        points for $\gamma=0.5$.}
        \label{fig:vis_dmr_cut}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
    \begin{figure}[]
        \centering
        \includegraphics[width=1.0\columnwidth]{monomer_cor_log.pdf}
        \caption{Monomer correlations in the QDPM.}
        \label{fig:vis_dmr_cut}
    \end{figure}
%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%
